### 2021-04-27 1.08-advisorless.Rmd

# Advisor-less {#advisorless}

During your PhD you will have had an advisor who would have been central to many of the decisions that you made. However your relationship with this person ended up, it is likely that you will still need advice from someone more experienced. If you are a postdoc, you should still have an academic whose research group you are working in, but as an early career researcher you may feel that you're suddenly cut adrift without advice. Increasing numbers of institutions are now recognising that early career researchers need mentors to advise them about how to move forwards within their academic field.

## What to look for in a mentor?

Doesn't have to be in the same institution, or even in the same country. In these days of post-COVID communication we are all a lot happier hooking up to meetings online.

## Questions to put to your mentor?

Where to publish / What meeting to go to / networks