### 2021-04-27 1.10-review.Rmd

# Writing a literature review {#review}

The first chapter of many PhDs is often a place for a literature review. Writing a review of the literature is actually a very useful task to undertake during your PhD, preferably towards the beginning. Firstly it will get you familiar with all of the relevant literature in your field. It will give you some idea of the historical context of your subject area. And it's especially good to help you spot gaps in the literature that may not have been realised before.

Reviewing the literature should make you familiar with the bigger questions in your subject area. This will make it easier for you to introduce each of your subsequent chapters.

## Having a purpose
Writing a literature review is easiest when you have a clear purpose. The objective can be particularly tight, or can aim to use the literature in order to answer a broader question. Either way having a clear purpose for your literature review will help you to define exactly what literature you're looking for.

Without an aim a literature review can be extremely daunting. There will be literally no end to the literature that you could include and it may well overwhelm you. If your question is more broad it will help to define a clear time period in which you are searching for literature. For example you may want to just consider the last 10 years. Remember that the amount of literature is increasing almost exponentially and so if you are covering the same ground as a previous review a lot can happen in 10 years.

If your aim is quite narrow then you can legitimately cover all of the literature. This may require delving back more than 100 years. But as we know from previous chapters the first journal only began around 150 years ago.

If you want your review to have an objective or statistical angle on the literature then consider writing a systematic review or a meta-analysis.

## Systematic reviews
I prefer to conduct a systematic review. This is where you define your search criteria and the databases that you want to use and then include all of the literature that comes out. You will need to provide clear criteria for the inclusion of literature. As well as explaining exactly how you decide to include or exclude papers that result from your search terms. Systematic reviews require a flow diagram to show how this literature was included or excluded. This means that you have to be systematic in how you record weather a paper is included or excluded.

I'm my opinion, a systematic review is less subjective than a review where the author is essentially cherry picking the literature in order to tell the story as they see it. It also allows you to conduct statistical tests in order to answer some of your questions (a meta-analysis). Lastly, it avoids one of the oldest problems in science: confirmation bias: Simply looking for articles that back up your own view. Too many old reviews are really just written by academics seeking to confirm their standpoints, and so moving forwards we must place the emphasis on being objective through systematic reviews. This view is now growing so that increasing numbers of journals will only accept a review if it is systematic.

## Using Google Scholar
Google scholar has distinct advantages over other literature databases as it searches words inside the contents of an article, not just in the title, abstract and keywords (like Scopus and WoS). However, the output style of Google Scholar is infuriating as it resists being able to capture search results into a simple table (database) format. In addition, it does not support the boolean search operators (see part 2), and has a far broader range of literature that is included.

Help is at hand for those of you who would like to use Google Scholar and get a database output with the publication of [Publish or Perish software](https://harzing.com/resources/publish-or-perish) (Harzing 2007). The original purpose of this software was to calculate personal research metrics for promotion purposes, and it still does all this. But it also provides a single platform for searching many different literature databases: Scopus, Web of Science, Microsoft Academic, Google Scholar.

Google Scholar has an important attribute over many other databases in that it returns results in languages other than English. Using only English to provide the basis for your review might lead you to draw an unreasonable conclusion with a literature bias (see Nuñez & Amano 2021). Language is only one of several potential biases in a literature review (see Amano & Sutherland 2013).

## The meta-analysis
A meta-analysis is a type of systematic review that uses the results of all of the studies included in order to provide a synthesis. While more powerful than a systematic review alone they take an awful lot of work. If you are expecting to get a lot of literature then consider putting a team together in order to conduct a meta-analysis.

Once written they can be very rewarding as they tend to attract a lot of attention and citations.

### Combining topics by combining people
The literature review that you conduct for your PhD will likely only have you doing the bulk of the work. But reviews become especially insightful when you combine topics together. To do this you may want to find somebody else who is an expert on the literature in that field. This might be somebody in your lab group. It might be somebody who is also just starting a PhD and is writing a literature review.

This brings me to another very important point. Share your enthusiasm for your topic. Remember to talk to people around you both in your laboratory and in your department about the work that you're doing, and tell them about the interesting things that you're finding. If they also tell you about their work then you may find that you have somebody to combine your literature review with in a novel angle that's never been thought of before. The paper that results could be a smash hit.

## Don't try to do too much
Whatever your aims were before you started compiling the literature, remember to remain flexible as you proceed. it's hard to know exactly how many papers you will encounter on a particular subject unless you already know it very well. If you're prepared to remain flexible while conducting your review it may save you from overreaching. Always be open to reducing the scope of your review or seeing particular questions or issues while you're reading that may be better than your initial idea.

## In summary
if you come to the end of your PhD and have not yet done a literature review then now is the time. You should already know a large chunk of the literature in your particular area and you will be looking for future questions to tackle in a postdoc. Conducting a literature review at this stage in your career will not only help you to highlight unanswered questions and interesting topics, but it will get your name out there. Once you've done it you'll be glad you have.