### 2021-04-27 2.02-letter.Rmd

# Letter to the editor {#letter}

Many journals require that you write a letter to the editor on submission. Each journal should tell you what this letter should contain. This will vary from journal to journal as some may have a set of check boxes that cover some points, while others may want a written declaration.

Here I provide a generic letter covering points that I usually cover in such letters. As with many things in this book, you should check the instructions to authors section.