### 2021-04-27 3.00-part3.Rmd

# (PART) Part 3: After your paper is accepted {-}

# Now that your manuscript has been accepted {-}

This part of the book gives you some things that you might want to do after your paper is accepted for publication. Much of this chapter is concerned with publicising the results and content beyond academia and for non-academic audiences. Even if you feel that there would be no interest beyond your academic niche, it would still be worth making some effort to popularise your study.