### 2021-04-27 4.04-paywall.Rmd

# Why did some journals go behind paywalls? {#paywall2}

Academics are not (usually) superstars, nor looking for enormous numbers of readers, but there would be little point to recording our work if we had no readers, or if our work were inaccessible, and so publishing is a necessity. However, we have got into a state in which much of our work is behind a paywall, and thus inaccessible to most. Whether or not we need our work to look pretty and appealing speaks more to our readers as humans than academics. Perhaps it goes without saying that an audience is likely to be larger the more appealing the presentation, and that’s not just the writing style but the layout and presentation of the text itself. And this is not new.

The first scientific journals appeared in 1665: Journal des Sçavans, and three months later Philosophical Transactions of the Royal Society of London. It is clear that the papers were type set and presented in the manner of a book, perhaps analogous to a collection of short stories. At the start, these were reports of studies that were presented at meetings. Producing proceedings of learned societies became the way in which most scientific journals began. Only later did it become possible to submit a manuscript that had not been presented. And later still when publishers began to manufacture their own scholarly journals in the absence of any academic society.

(ref:scavans) The first issues of the World's first scientific journals

```{r scavans, echo=FALSE, out.width = '50%',  fig.cap="(ref:scavans)"}
knitr::include_graphics('figures/scavans.png')
```

(ref:PhilTransRoySoc) Note that the Journal des Sçavans ceased publication in 1792 due to the French Revolution, but that the Philosophical Transactions of the Royal Society of London continues to this day as the world’s longest running scientific journal.

```{r PhilTransRoySoc, echo=FALSE, out.width = '50%',  fig.cap="(ref:PhilTransRoySoc)"}
knitr::include_graphics('figures/PhilTransRoySoc.png')
```

Being the editor of a society journal means being elected by members of that society, and being responsible to an editorial board, normally made up of the society’s members. Until very recently, and I’m thinking back to my first interactions with editors for my first few publications, submitting to the journal meant producing three (or sometimes more) double spaced copies of a manuscript and mailing them to the editor. Editors of bigger journals had secretaries dedicated to handling the administration of the paper. Following a telephonic enquiry, the copies were sent out to referees by post and sent back to the editorial office with a typed report often together with the marked up manuscript. Once the editor had received all reports, they communicated their decision back to the corresponding author (i.e. the one to whom correspondence was addressed) and, once accepted, the article went into production. Prior to personal computers being commonplace (only 25 years ago), each journal would have had to have had a publisher to set the type and print the pages. Clearly, this was beyond the scope of individual societies and the publisher was a necessity. Libraries had to pay for copies of journals, as the cost of publishing had to be offset by the society.

The advent of desktop publishing changed the need for publishers and brought many small society journals onto a larger platform where they could produce attractive content (over the typewritten documents that had been stencil duplicated or similar) and sent around to members. However, for small societies there was no professionalism involved as it devolved to the editor to publish society material. This is where I entered the stage in 2009 when I took over as editor of the _African Journal of Herpetology_. Thankfully, email had taken the place of the postal service, but once a manuscript was accepted I was the one who needed to type-set the documents (in Quark) and send out proofs to authors. Once proofs were corrected and the issue was ready, I had to find quotes from 3 printers, deliver discs and ultimately collect boxes of printed journals. Back at home, I also packed all of the copies into labelled envelopes [(with some help from friends)](https://youtu.be/sJReMVz7Yxk) and carried the boxes once again to the post office.

(ref:journal-video) Stuffing envelopes with jouranl copies seemed like an endless task. View [on YouTube](https://youtu.be/sJReMVz7Yxk).

```{r journal-video, echo=FALSE, fig.cap="(ref:journal-video)"}
if(knitr::is_html_output(excludes="markdown")) knitr::include_url("https://www.youtube.com/embed/sJReMVz7Yxk")
```

After the first issue, I realised that I could not do all of this work indefinitely. I knew that there were publishers who were interested in acquiring the journal into their stable, and I contacted them and started negotiations. In 2011, the first copy of AJH from a professional publisher emerged, and allowed me to go back to editing the content through peer review.

At the time, I was aware of Open Access and considered this as an option for the journal. Open Access would have required that someone pay for the type setting, and the society would still have to pay for an online dissemination platform, given that they did not have their own platform. This would have meant that authors paid for getting their manuscripts into print. And then, like today, the decision was that our authors would not be willing to pay. Other, richer, societies were able to go Open Access with the costs being covered by the authors. For some this became an incredibly successful model, with submissions increasing as well as the Impact Factor. They demonstrate that Open Access is possible on an independent platform.

## Why don’t all society journals do without publishers, and go Open Access independently?
The first problem is that societies generate income from journals. Subscribers to print or virtual copies pay the society, and this can defray a large part of the cost of publishing otherwise carried by the members. Some make a profit, and this profit can enable the society to do more for their community of members. This could include providing small grants, subsidising conferences for students and other much appreciated initiatives. Going Open Access means losing this revenue, as well as taking on the extra costs.

The second problem for many societies is that their members are paying, and the councils or committees elected to represent the members do not feel that it is fair for their members to pay for open access for everyone else. Part of the privilege of being a member of the society is having a free (or more accurately paid through membership) copy of the society’s journals. The costs are not high, and the exclusivity of members having Open Access while it is denied to others is perhaps just a hangover from the days when the only other copies were in the library. Certainly, the costs are nothing like those which authors are now charged by publishers to turn their accepted manuscripts into Open Access.

But the current situation is unsustainable. Tax-payers (in the main) pay for science to be conducted at universities and other institutes. In places where Open Access is mandatory, the tax-payer pays again to have the research published at a cost that is far inflated from the actual production costs. Publishers are getting fat, and the losers are the scientists (whose funding is reduced to pay publication costs) and the tax-payers who end up feeding the greedy publishers.

## What will it take to break the vicious cycle?
We need new models for publishing. Society journals are still kings in this game and ultimately hold the cards for moving away from filling the pockets of publishing companies. What we have seen in recent years is that journals can come from nowhere to become dominant players in the system. Think PLoS ONE, and the even more recent Scientific Reports. These mega Open Access journals didn’t exist 10 years ago. And they don’t need to exist 10 years from now. What is needed is for the actual costs of publishing (not that currently inflated by publishers) to be covered by the institutions that employ the academics. This could cover type-setting and the additional IT infrastructure (on-line submission system and online dissemination platform). Most (if not all) societies are not-for-profit organisations, and only need to cover the costs of publishing.

## Can we afford not to change?
If you are from a rich country or institution, then you can probably afford the current system. Those who cannot are researchers in disadvantaged countries. In some cases, the cost of publishing Open Access is greater than the cost of the research. These are insurmountable costs for many researchers. We have a massive hole in scientific contributions from the poorest of nations, and the current Open Access models will see their work being the most hidden from view, while the countries paying for their work do so disproportionately. But even developing countries could be winners in a new Open Access model. By sourcing the relevant IT skills in the country, governments of middle-income countries could facilitate the content of their own society’s with relative comfort. In my opinion, everyone should publish work from scientists in the poorest countries as Open Access without any charges.

## Societies need money. Editors can’t be publishers.
We need free software. We desperately need good, free editorial management software. There are some free versions out there, but what we need are free versions that are at least as good, if not better, than existing platforms (e.g. ScholarOne; Editorial Manager). Galipeau et al [-@galipeau2016scoping] make the point about the ever expanding role of editors in the modern publishing era. There is no scope for editors to take on extra work.

We need a free La-Tex interface with robust templates for all societal journals. Ideally, this would be packaged with the above editorial management software. This must have the ability to cope with figures and equations, and the unusual demands that some society journals have.

We need a solution for hosting and disseminating the Open Access society journals (and their supplementary information if not hosted elsewhere) in perpetuity. This last point is perhaps the most expensive, and almost certainly requires government assistance. Maybe this is an interesting use of the block-chain with libraries keeping the data. It would be an interesting way to build a doi with editor and referee unique IDs, and the document's information hanging off.

We need to take back our content stuck behind paywalls. Yes, it’s time for you all to dig up those old submitted manuscripts and submit them to an institutional repository where it can be accessed for free.

## A paywall is never acceptable wherever you put it
Any paywall, whether it be high (i.e. EUR 9500 for Nature in 2020), or ‘considerably lower’ (i.e. EUR 900 for NeoBiota in 2020) is a wall that excludes many researchers, and certainly those from many developing countries. I have made this point before [@measey2018europe]. While previously I lobbied for the source of these publishing fees to become public - which would show that for a great majority of researchers, publishing fees are coming from research funds. Funds that would otherwise further knowledge are going directly into the pockets of publishers. Publishers instantly refused to do this. Now I think that our energies would be better spent demolishing the paywall (what they refer to as Diamond Open Access). And for almost all of us, this means doing so without the benefit of a rich uncle.

Until we have Diamond Open Access for all, having the paywall after publication is actually preferable for most of us, as most of us cannot afford to pay anything as we simply don’t have this type of discretionary funds. We do have to publish our work, and would rather that it was out there behind a paywall, than not out there at all.

I am not advocating a paywall, but I disagree that by placing the paywall before publication (i.e. on acceptance) solves anything for anyone other than the most privileged researchers. In the words of Peterson et al [-@peterson2019open], "do not replace one problem with another". Instead, what we need is to tear down the paywall with a completely new publishing model for academia. We are all quick and ready to agree that Open Access is the best way forward for all scientific results. The aims and objectives of Science Europe are laudable and will lead to a far better publishing environment for European scientists. Scientists from other high-income countries will also benefit from this decision, having a far greater number of journals in which to publish Open Access.

If current trends continue, scientists from low-income countries will be granted full fee waivers. Many journals use the [Hinari Eligibility list of countries](http://www.who.int/hinari/eligibility/en/) to separate Group A (free access) and Group B (low cost access – normally billed at a 50% reduction in fees). The lists are made up from five global economic and development criteria.

Middle-income countries are missing from these lists, and receive no support for fee waivers. Their governments provide scientists with no means of paying fees. Scientists who pay fees often do so from their own research budgets. The increasing number of journals that charge unjustifiably high publishing costs are forcing middle-income scientists away from Open Access journals. In my own lab, publication fees are regularly more than the cost of conducting much of our ecological research.

## The answer lies inside our University Libraries 
The university library has undergone a massive transformation over the last 20 years. During my PhD, I made a weekly visit to the library to physically pick up the latest issues of all the journals that came through the postal service from all over the world. For papers that I found out about but had no access to, I had a stack of postcards that were specifically for reprint requests, and I enthusiastically filled them out and posted them off to researchers the world over. Librarians arranged these issues on the shelves and eventually sent them off for binding into volumes and then worried about the physical space that was available inside the library as every year publication inflation [@lariviere2016how] meant more pages to be supported within their walls.

Probably the most stressful time in the library now is around negotiating the next contract with a mega-publisher. Will they be able to meet next year’s demand for cash? How much are other universities paying? Of course, the bundles are sold with non-disclosure agreements, so that librarians who successfully negotiate a lower price at one institution cannot influence the negotiations at another. Doesn’t this sound like an extortion racket?

The solution then would be for our learned academic societies to come to agreements directly with university libraries. Today, publishing is less about type-setting (which many top publishers outsource and do very poorly), and more about dissemination. This is something that our library and academic librarians have been doing for decades, and do far better and far more cheaply than any publishing house. Given the choice, most of us would prefer to entrust our academic endeavors to our own libraries rather than to for-profit publishing houses.

There are more reasons why it makes sense for libraries to take on the roll as publishers. Most of us are employed by universities or research institutions that also fund our libraries. Linking the work we do (writing, reviewing and editing) more closely with our institutions would result in a greater appreciation for this part of our workload. Editors and associate editors will appreciate that they get little credit from their institutions for a considerable extra amount of work that they perform.

Libraries have fantastic networks, and are our professional long-term storage partners. They developed efficient and impressive information technology (IT) long before it hit most academic departments. Their inter-library networks are what we now need to disseminate the knowledge that we generate without any walls.

The idea of libraries as the new publishers isn't new. Raju and Pietersen [@raju2017library] proposed this as a solution in Africa. Here I extend the same idea as an exclusive way of publishing academic journals for the world.

## We need to give up our addiction to fancy layouts
Once the storage and dissemination of our contributions are taken care of, the only service left from the publishers is a fancy layout. This is mostly a historical legacy which I've talked more about in the past (see above). I have to admit that I really like seeing my work being nicely produced and printed. But I’m happy to give this up if it means demolishing paywalls. In reality, LaTex can solve most of these problems so that we simply use the journal (library) produced template, that will need minimal manipulation afterwards.

We also need to persuade our employers to stop their obsession with publishing metrics, for example by adopting the [San Francisco Declaration on Research Assessment](https://sfdora.org/read/) (known as DORA).

No doubt, there will be some institutions that will invest extra to have nicer layouts. But I feel confident that this will not change the impact factor, or any other metrics, as academics will value the content for what it contains rather than what it looks like. Admittedly, nothing about the contents of the highest ranking journals suggests that impact factor is consistently related to research quality.

If you have read this far, then I hope that you will join the call for Diamond Open Access - no paywalls for anyone.