### 2021-04-27 5.00-lastnote.Rmd

# Last note {- #lastnote}

I really hope that this book has been helpful and that it has achieved what I set out to do: provide you with the guide on how to write a PhD in biological sciences. If you feel that this book has important items missing, out-of-date or simply wrong, then please help. Any good guide relies upon the people that use it to keep it viable.

You can contribute to the project at the Github pages for this book, using bookdown.

If I could emphasise one really important aspect of all of your PhD studies; that is to enjoy them!