### 2021-07-31 4.09-mentalhealth.Rmd

# Keeping track of your mental health {#mentalhealth}

Stress is a natural part of life, and many people are at their most productive when they are under some degree of pressure, such as a deadline. Although deadlines don’t work for everyone. Problems arise when we become overwhelmed by stress and are unable to fully respond. When this occurs productivity can drop off and survival responses can be triggered as if responding to an actual physical attack. These responses include fight, flight or freeze responses. Anxiety and panic can be triggered. In this state additional demands on your time may also push your life off balance, so that you start to neglect your personal wellbeing which can negatively impact on relationships, exercise regime, or even nutrition and personal hygiene. Some people can find that the additional stress can cause physical symptoms that may even need medical treatment. Your sense of competence and mastery can be negatively impacted such that you may even suffer from feelings of inadequacy or imposter syndrome. Impostor syndrome is an experience you have when you believe that you are not as competent as you think others perceive you to be. It is not uncommon in many professions, and especially prevalent in academia [@clance1978imposter]. This is now widely recognised and there are lots of useful shared experience out there to read [e.g. @dickerson2019how].

The General Health Questionnaire is an instrument used to measure psychological distress. For example, you could use the GHQ-12 in [Table 4.1](http://www.howtowriteaphd.org/healthy.html#tab:GHQ) [see @measey2021how]. It is quick, reliable and simple to score, so you can use it at any time during your career as an indicator of whether you need to reach out to personal, occupational or professional support networks. My suggestion is that you complete a General Health Questionnaire now and record your answers as a baseline. Keep the scores somewhere safe. If you feel that you are getting overly stressed, take the test again and compare them with your baseline scores. Although there are no hard rules, if three or more of your scores have moved by two or more points it could be worth discussing with your support network to help you decide whether or not to seek professional help.

Even if you don't feel you need the support of your institution now, it is worth finding out how they can support your mental health in the future if needed. Although there has been some stigma attached to difficulties with mental health in the past, most institutions accept that pressures are mounting on their staff and that they may require support. Most institutions will make experienced councillors available to support you if needed. Importantly, you should realise that none of these symptoms are unusual and that there is a high probability that many of your colleagues may also be struggling. Knowing that your problems are shared and reaching out to support networks early is an excellent way to prevent them from escalating beyond your control. Your best means of coping will be to try and develop a support network and to understand where and with whom you can discuss any difficulties as they arise. Knowing who this is and how and when to approach them will put you in a stronger position if you need them in future.