### 2023-06-13 index.Rmd
Add files via upload
# fixing McCoy suggestions for Conflicts of Interest

--- 
title: "How to publish in Biological Sciences: a guide for the uninitiated"
author: "John Measey"
date: "`r Sys.Date()`"
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
csl: peerj.csl
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
site: bookdown::bookdown_site
#output: bookdown::gitbook
description: "Publishing a paper in an academic journal should simply consist of submitting a publication worthy manuscript. But having a working knowledge of publishing will enable you to make better decisions about what, where and how to submit manuscripts. This all comes with experience, and in this book, I unravel areas of assumed knowledge and furnish them with explanations pitched at the Early Career Researcher, along with links and citations where you can read more. This book explores the wicked problem of academic publishing, and how to move from a closed to an open, transparent system. I explain the many choices that exist for those wishing to submit a manuscript for publication in the Biological Sciences. I explore the world of publication bias, and how this is evidenced by reviewers and editors. Impact Factors, citations and how the desire to track the performance of academics has led to unethical practices and exploitation of science and scientists. This guide provides an 'everything you wanted to know about publishing but were afraid to ask' approach for anyone who still feels that submitting a manuscript is like posting it into a black box. This book is written to get you onto an even footing."
github-repo: johnmeasey/How-to-publish-in-Biological-Sciences
graphics: yes
cover-image: figures/cover.png
favicon: figures/favicon.png
---

```{r setup, include=FALSE}
options(
  htmltools.dir.version = FALSE, formatR.indent = 2, width = 55, digits = 4
)

# install the packages needed by this book; you fill out c(), e.g. c('ggplot2', 'dplyr')
lapply(c('xfun'), function(pkg) {
  if (system.file(package = pkg) == '') install.packages(pkg)
})
```

# How to publish in Biological Sciences  {-}

This book was published in **2022** with CRC Press. The printed version of this book is available from [CRC Press](https://www.routledge.com/How-to-Publish-in-Biological-Sciences-A-Guide-for-the-Uninitiated/Measey/p/book/9781032116419). Citation details can be found [here](#how-to-cite).

```{r fig.align='right', echo=FALSE, include=identical(knitr:::pandoc_to(), 'html'), fig.link='http://john.measey.com/'}
knitr::include_graphics('figures/cover.jpg', dpi = NA)
```

You can order your own copy on the [CRC Press website](https://www.routledge.com/How-to-Publish-in-Biological-Sciences-A-Guide-for-the-Uninitiated/Measey/p/book/9781032116419), [Amazon](https://www.amazon.com/How-Publish-Biological-Sciences-Uninitiated/dp/1032116447), [Barnes & Noble](https://www.barnesandnoble.com/w/how-to-publish-in-biological-sciences-john-measey/1140172276) and other good book sellers. Please note that the printed version of this book dates back to July 2022 and that the online version continues to grow and change. Take a look [here](#contribute).

**Cover illustration**: The Tower of Babel was adapted by the author from an original work by Pieter Bruegel the Elder (1563) painted to represent the demolition and scattering of peoples so as to impair their communication. The collapse is coming, but can we save the scientific project? In my version, a group of Early Career Researchers can be seen (bottom left) deciding how to rebuild the ancient edifice that represents scientific publishing (centre), with an entrance guarded by gatekeepers (centre right). The skies above reflect the choice open to them; to stay in a closed system or to choose a future of open science.

# Welcome {-}
<br>
Welcome to my guide on how to publish in the biological sciences. This guide is pitched at the early career researcher. It is not going to tell you what to write, but to open and examine the black box of scientific publishing, and more broadly explore how this impacts the academic context. My intention is to demystify publication in the biological sciences, so that readers become aware of what is happening once they have submitted a manuscript, and how to better interpret the decisions made by colleagues who are reviewing and editing your work.

Publishing has become vital for all academics, such that it is widely recognised that we inhabit a 'publish or perish' academic landscape. For some the process appears effortless, while for others publishing represents a black box leaving them outside in a highly stressful environment. This book is meant to be a guide to those uninitiated members of the academic community, postgraduate students and early career researchers, to bring them up to speed with all the necessary background information on publishing, providing links and references for reading and learning more.

## Why read this book {-}

Publishing a paper in an academic journal should simply consist of submitting a publication worthy manuscript. But having a working knowledge of publishing will enable you to make better decisions about what, where and how to submit manuscripts. This all comes with experience, and in this book, I try to explore the areas of assumed knowledge, and furnish them with explanations pitched at the Early Career Researcher, along with links and citations where you can read more. I explain the many choices that exist for those wishing to submit a manuscript for publication in the Biological Sciences. Where possible, information in this book are tailored for the biological sciences, and when this information was not available for the Life Sciences, STEM or science (in that order). I explore the world of publication bias, and how this is evidenced by reviewers and editors. In many cases, Impact Factors, citations and how the desire to track the performance of academics has led to unethical practices and exploitation of science and scientists. This guide provides an "everything you wanted to know about publishing but were afraid to ask" approach for anyone who still feels that submitting a manuscript is like posting it into a black box. This book is written to get you onto an even footing.

## What's not in this book {-}

Depending on just how early you are in your career, there is a lot missing from this book that has been provided in another book [**How to write a PhD in Biological Sciences**](https://howtowriteaphd.org/). That book concentrates on getting PhD students started writing data chapters, while this book concentrates on publishing manuscripts. Hence, if you want extra information about writing in the biological sciences, I would point you to the other book. If you are happy with what you have written, but want help to demystify the publishing process, then this is the right book for you.

If you want more help with writing the manuscript, then [**How to write a PhD in Biological Sciences**](https://howtowriteaphd.org/) is the content that you need:

```{r cover2, echo=FALSE, fig.align='centre', out.width='50%', fig.link='https://howtowriteaphd.org/'}
knitr::include_graphics('figures/Cover1.jpg')
```

There is common ground in both books, and I will point to important chapters relevant to publishing in the other book that are not reproduced here:
[Plagiarism](http://www.howtowriteaphd.org/plagiarise.html), [Being aware that you can get it wrong](http://www.howtowriteaphd.org/TypeI.html), [Transparency](http://www.howtowriteaphd.org/transparency.html), and many more.

## Structure of this book {-}

This book is written in four parts:

[**Part I - Getting your manuscript ready for submission**](#part1)
Although you may have already done your research, and written your manuscript, getting it ready for publication will require a new set of hurdles for you to jump over. In this section, I discuss what you need to know before entering into the publication arena. What are scientific journals for, and who are the gatekeepers? How does peer review work? The publishing world is at a turning point, and before you start publishing you should be aware of the current reality in Biological Sciences around the currency of citations and how these relate to other metrics such as the Impact Factor and career advancement. You also need to know potential directions for publishing, including the need for transparency in your work, whether or not you should deposit your manuscript as a preprint, and who you should invite to be an author. Tactical chapters for Early Career Researchers provide information on how to actively build and maintain a network to facilitate and support your work.

[**Part II - Submission, reviews and reviewing, revising and resubmitting**](#part2)
Sending a paper to a journal is like posting it into a black box where, after some time, you might simply get a rejection and have no idea what has happened. In this section, I take you through the mechanisms of submitting a manuscript from choosing the right journal for your submission, writing a letter to the editor, suggesting reviewers, entering metadata about your manuscript into the editorial management software, all the way to pressing the submit button. I explain how the editorial submission system works, and what you can expect from editors and peer reviewers. I take a practical approach to writing a rebuttal and explain how and why you should expect to revise your manuscript for the editor. The eventual goal of this section is to demystify the entire process between submission and acceptance, and to understand the process from the viewpoint of an author, editor and reviewer.

[**Part III - Once your paper is published**](#part3)
Once your article is accepted, you can celebrate together with your co-authors! You will need to submit the final version of your manuscript, have this type-set and then approve the proofs before a Version of Record appears. At this point, you can start to share your paper, but there are still some key steps that you can take to improve the dissemination of the research both to the academic community, to your funders, and the public at large. Who is it best to share your research with, and what would be the best form to share it in? In this section, there are chapters that explain how to write a press release and a popular article on your paper, and how you can improve and monitor its circulation both in academia and in the general media and social-media, focussing on those stakeholders who might use your findings.

[**Part IV - Further challenges in academia**](#part4)
 The last part of this book discusses the wicked problem posed by current publication models in academia. This section deals with the growing number of issues driven by a 'publish or perish' culture, and what this means for Early Career Researchers. Special focus is given to the paywall erected by many publishers, Open Access publishing and predatory publishers. I also explain the problems in the current system of biases in peer review, and the confirmation bias in scientific publishing. Instead of just presenting you with problems, this section provides insight into ideas that the academic community has produced in order to get over the current problems. Other important hurdles that you might meet, such as retractions, fraud and bullying, receive in depth focus.

## Why 'A guide for the uninitiated'? {-}
Early Career Researchers are within eight years of getting their PhD or within six years of their first academic job. At this time, you will have already experienced the academic life, including publishing, but there will be far more to it than you are aware of. This book considers Early Career Researchers as colleagues who simply lack the experience of a system that has changed in many ways over the last 20 years. To those of you who know the current publishing scene in biological sciences, it offers the perspective of where things have come from. I have written this book as I feel that I would have been able to achieve more had I understood more about the publishing process early on in my own career. If I had only had a guide to tell me what it was all about, I could have saved myself so much stress, time and energy. In short, I feel that I was uninitiated, and this is the guide I wish that I had had. So, this guide is my practical attempt to help you; to get you up to speed in the world of academic publishing, specifically in the biological sciences. After reading this book, I hope that you will avoid the nightmare world of publishing of constant effort ending in dead end rejections that so many academics describe.

<br>

## Acknowledgments {-}

There are a great many people that I need to thank. First and foremost are the members of my lab, past and present, who have inspired me to put together first the blog posts and then the book. It is because you wanted more that I put this together. This book contains lots of links to blogs and articles written and posted freely on the internet by others who also aim to demystify and help. I thank this greater academic community (especially \#academicTwitter) for sharing and inspiring. Thanks go to the many reviewers and editors who have taken their time to improve my writing. I am still learning. Lots of the text in this book has been improved by feedback from my students and postdocs. A special mention must go to my brother, Richard, who has hosted my lab website for more than a decade, and especially for saving blog posts from hacking attacks. Thanks also to Thalassa Matthews, who proofread many of the blog posts after I had published them late at night, so that I could correct them over breakfast in the morning. Graham Alexander, James Baxter-Gilbert, Dan Bolnick, Chris Chambers, Jack Dougherty, João Fontiela, Thomas Guillemaud, Anthony Herrel, Michael Hochberg, Andrea Hurst, Allan Ellis, Rachael Lammey (CrossRef), Krista McCoy (and the McCoy lab group), Andrea Melotto, Lisa and Mark O'Connell, Ivan Oransky (Retraction Watch), Heather Piwowar (ImpactStory), Claire Riss (Center for Open Science), Johan Rooryck (cOAltition S), Daniel Shea, Daniel van Blerk, James Vonesh, Carla Wagener all read or commented on different aspects of the book. Thanks are also due to my colleagues at the Centre for Invasion Biology, the Department of Botany and Zoology, and Stellenbosch University. A special thanks to the librarians who have supported many of my more extreme rantings about publishers.

```{block2, type='flushright', html.tag='p'}
John Measey  
Cape Town
```